---
title: "HW2"
author: "Corey Kuhn"
date: "2/7/2018"
output: pdf_document
---

\newcommand{\newprob}
    {
    \vskip 1cm
    }
\newcommand{\newln}
    {
    \vskip .1cm
    }

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```


## Problem 1 - Chapter 5, Exercise 3

**Part(a)**

The following function simulates a game of Chevalier de Mere, where you are betting that you get at least one 6 in four throws of a die. The input of the function $x$ is the number of throws (default is 4 for this particular game). The function outputs the outcome of each roll and whether you win or lose.  
\newln
```{r}
CdM <- function(x=4){
  rolls <- sample(1:6,size=x,replace=TRUE)
  outcome <- ifelse(6 %in% rolls, "WIN", "LOSE")
  return(c(rolls,outcome))
}
# Examples
set.seed(12)
CdM()
CdM()
```

\newprob

**Part(b)**

Same function as in Part (a) with only a few minor details changed. By default, n=4, as defined in the argument of the function.  
\newln
```{r}
sixes <- function(n=4){
  rolls <- sample(1:6,size=n,replace=TRUE)
  outcome <- ifelse(6 %in% rolls, "WIN", "LOSE")
  return(c(outcome))
}
# Examples
set.seed(12)
sixes()
sixes(7)
```

\newprob

**Part(c)**

The code below conducts 100 runs for each value of N and takes a summary of the proportions of wins and the differences between the simulated and theoretical probabilities of wins. Looking at the first set of summaries, we see that for N=100 the proportion of wins varies from a minumum of 0.43 to a maximum of 0.64, for N=1000 the proportion of wins varies from a minumum of 0.4810 to a maximum of 0.5630, and for N=10000 the proportion of wins varies from a minumum of 0.5036 to a maximum of 0.5345. We can see that as N increases, the variability in the proportion of wins decreases, since the range of proportions becomes tighter.
\newln
Looking at the second set of summaries, we see that the differences between the simulated and theoretical probabilities decrease as N increases from 100 to 10000.
\newln
```{r}
# prop.win() determines the proportion of wins in N simulations of the game.
prop.win <- function(N, n=4){
  sim <- replicate(N, sixes())
  prop.win <- length(sim[sim=="WIN"])/length(sim)
  theo.win <- 1-(5/6)^n
  dif <- abs(prop.win-theo.win)
  return(c(prop.win, theo.win, dif))
}
# Conduct 100 runs for each value of N and find the summary of proportions of wins
runs <- 100
set.seed(12)
summary(replicate(runs, prop.win(100))[1,]) # N=100
summary(replicate(runs, prop.win(1000))[1,]) # N=1000
summary(replicate(runs, prop.win(10000))[1,]) # N=10000
# Find the summary of difference between simulation probability and theoretical probability
set.seed(12)
summary(replicate(runs, prop.win(100))[3,]) # N=100
summary(replicate(runs, prop.win(1000))[3,]) # N=1000
summary(replicate(runs, prop.win(10000))[3,]) # N=10000
```

\newprob

**Part(d)**

The `write.txt` function below creates a text file with the results of the game for N simulations. You must specify a value N. The `prop` function reads the text file containing the N simulations and calculates the proportion of wins. You must input a file path in quotes if different from `prop("~/Desktop/sixes_sim.txt")`.
\newln
```{r}
# Writes a file
write.txt <- function(N){
  write(replicate(N, sixes()),"~/Desktop/sixes_sim.txt")
}
# Reads a file and determines proportion of bets won. Must provide the file path/name for the function if different from file="~/Desktop/sixes_sim.txt".
prop <- function(file="~/Desktop/sixes_sim.txt"){
  sim <- read.table(file)[,1]
  prop.win <- length(sim[sim=="WIN"])/length(sim)
  return(prop.win)
}
```

\newpage

## Problem 2 - Chapter 5, Exercise 5

The function `dds` defined below outputs the first $n$ terms in the discrete dynamical system defined by $f$, and then plots them. There are 3 inputs to the function: `x1` is the first value in the system and must be between 0 and 1, `r` is a value to be defined between 0 and 4, and `n` is the number of terms in the system to output.
\newln
```{r, fig.width=6, fig.height=6}
dds <- function(x1, r, n){
  if(x1 > 1 | x1 < 0){
    return("Error: x1 must be between 0 and 1")
  }
  sequence <- c(x1,rep(NA,n-1))
  for (i in 2:n) {
    sequence[i] <- r*sequence[i-1]*(1-sequence[i-1])
  }
  plot(1:n,sequence, xlab="n", ylab="x[n]", main=paste("r = ", r, ",", "x1 =", x1))
  return(sequence)
}
# Example with x1=0.6, n=100, and r taking on various values defined in vector below
r <- c(1.5,2.9,3.1,3.5,3.56,3.57,3.58,3.8,4)
par(mfrow=c(3,3))
for (i in r){
  dds(0.6, i, 100)
}
# Example with x1=0.2, n and r remaining unchanged
par(mfrow=c(3,3))
for (i in r){
  dds(0.2, i, 100)
}
```

\newpage

## Problem 3

```{r, message=F, warning=F}
# Histogram using R base graphics
hist(mtcars$mpg)
# Histogram using ggplot2
library(ggplot2)
ggplot(data=mtcars, aes(x=mpg)) + geom_histogram() + ggtitle("Mtcars mpg")
```

\newpage

## Problem 4

```{r}
# Scatterplot using R base graphics
plot(mtcars$wt, mtcars$mpg, main="Mpg vs. Weight")
# Scatterplot using ggplot2
ggplot(data=mtcars,aes(x=wt,y=mpg)) + geom_point() + ggtitle("Mpg vs. Weight")
```

\newpage

## Problem 5

```{r, message=F, warning=F}
# Creating the first plot
mtcars$am <- as.factor(mtcars$am)
levels(mtcars$am) <- c("Automatic","Manual")
ggplot(data=mtcars,aes(x=factor(am),y=mpg))+
  geom_violin(aes(fill=factor(am)))+
  theme(axis.text.x = element_blank(),
        axis.ticks=element_blank())+
  ylab("Miles per Gallon (MPG)")+
  scale_fill_discrete(name="Type")+
  scale_x_discrete(name="")+
  ggtitle("MPG by transmission")
# Creating the second plot
library(dplyr)
ggplot(data=diamonds,aes(x=carat,y=price)) + geom_point() + geom_smooth() + facet_wrap(~cut)
```







