---
title: "HW4"
author: "Corey Kuhn"
date: "3/27/2018"
output: pdf_document
---

\newcommand{\newprob}
    {
    \vskip 1cm
    }
\newcommand{\newln}
    {
    \vskip .1cm
    }

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```


## Problem 1 - Chapter 18, Exercise 4  

cdf: $F(x) = P(X \leq x) = \sum_{x_i \leq x}^{}{f(x_i)}$  

\begin{tabular}{ c c c c c }
\hline
 x & 0 & 1 & 2 & 3 \\ 
 \hline
 F(x) & 0 & 0.1 & 0.4 & 1 \\
 \hline
\end{tabular}  

To simulate a RV with this distribution, use the two functions below. More specifically, doing `cdf.sim(cdf)` with return a value from the distribution defined by the pdf.
\newln
```{r}
# Plot cdf
plot(stepfun(1:5,c(0,0.1,0.4,0.4,0.4,1)), main="CDF")

# Simulating RV with this distribution
cdf <- function(x){
  if(x < 1){
    Fx <- 0
  } else if(x>=1 & x<2){
    Fx <- 0.1
  } else if(x>=2 & x<5){
    Fx <- 0.4
  } else {
    Fx <- 1
  }
  return(Fx)
}
cdf.sim <- function(F, ...){
  X <- 0
  U <- runif(1)
  while(F(X) < U){
    X <- X + 1
  }
  return(X)
}

# 1,000 simulations
nsim <- 1000
sims <- rep(NA,1000)
set.seed(1234)
for(i in 1:length(sims)){
  sims[i] <- cdf.sim(cdf)
}
table(sims) # Proportions are very close to the pdf probabilities
```


\newpage

## Problem 3 - Chapter 18, Exercise 13

$\mathbf{F_X(x)} = \int_1^x 2(x-1)^2 dx$  

Let $u=x-1$ and $du=dx$.  

$= \int_0^{x-1} 2u^2 du = \frac{2}{3}u^3 | _0^{x-1}$  

$\mathbf{= \frac{2}{3}(x-1)^3}$  

Now to find the inverse function $F_X^{-1}(y)$, set  

$y = \frac{2}{3}(x-1)^3$  

and solve for $x$:  

$x = 1 + (\frac{3}{2}y)^{1/3}$  

Thus,

$\mathbf{F_X^{-1}(y) = 1 + (\frac{3}{2}y)^{1/3}}$  

The pdf in this problem is not a valid pdf, since the integral evaluated from on the x limits from $x=1$ to $x=2$ does not equal 1. But if this was a valid pdf, the cdf is plotted below.  

We can simulate a random variable with the cdf using the inversion method by defining the inverse function $F_X^{-1}(y)$, randomly selecting from a uniform distribution, and plugging these random uniform draws into the inverse function to return the random draws from the particular distribution defined in this problem. Defined in the pdf in the problem, we see that $x$ only takes on values between 1 and 2, but looking at the summary of our simulations below, we see that there were values simulated up to 2.144. This may be because the pdf defined in the problem is not a valid pdf.
\newln
```{r}
# pdf
pdf <- function(x){
  fx <- 2*(x-1)^2
  return(fx)
}
integrate(pdf,1,2)$value # pdf integrates to 0.6667, not a valid pdf!

# Plot cdf
cdf <- function(x){
  Fx <- (2/3)*(x-1)^3
  return(Fx)
}
plot(seq(1,2,0.01), cdf(seq(1,2,0.01)),main="CDF", type="l")

# Simulate a RV with this cdf using inversion method
cdf_inv <- function(y){
  Fy <- 1 + ((3/2)*y)^(1/3)
  return(Fy)
}
set.seed(1234)
u <- runif(1000)
sims <- cdf_inv(u)
summary(sims)
```

\newpage

## Problem 3 - Chapter 18, Exercise 13

$\mathbf{F_X(x)} = \int_{-\infty}^x \frac{e^{-x}}{(1+e^{-x})^2} dx$  

Let $u=1+e^{-x}$ and $-du=e^{-x}dx$.  

$=\int\frac{-1}{u^2}du$  

$=-\frac{u^{-1}}{-1} = \frac{1}{1+e^{-x}} | _{-\infty}^x$  

$\mathbf{= \frac{1}{1+e^{-x}}}$  

Now to find the inverse function $F_X^{-1}(y)$, set  

$y = \frac{1}{1+e^{-x}}$  

and solve for $x$:  

$x = -log(\frac{1}{y} - 1)$  

Thus,

$\mathbf{F_X^{-1}(y) = -log(\frac{1}{y} - 1)}$  

See the below code for the cdf and for simulating draws from a standard logistic distribution.
\newln
```{r}
# Check to make sure pdf integrates to 1
pdf <- function(x){
  fx <- exp(-x)/(1+exp(-x))^2
  return(fx)
}
integrate(pdf,-20,20)$value # valid pdf!

# Find cdf
cdf <- function(x){
  Fx <- 1/(1+exp(-x))
  return(Fx)
}

# Define inverse cdf
cdf_inv <- function(y){
  Fy <- -log((1/y)-1)
  return(Fy)
}

# Drawing randomly from the distribution
set.seed(1234)
u <- runif(1000)
sims <- cdf_inv(u)
summary(sims)
```



