---
title: "HW5"
author: "Corey Kuhn"
date: "4/24/2018"
output: pdf_document
---

\newcommand{\newprob}
    {
    \vskip 1cm
    }
\newcommand{\newln}
    {
    \vskip .1cm
    }

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, warning=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

## Problem 1

The function `rand.letter()` returns a randomly chosen letter of the alphabet.
\newln
```{r}
rand.letter <- function(x){
  let <- sample(letters)[1]
  return(let)
}
```

\newprob

## Problem 2

Running the function 1 million times in parallel take about half the time it takes to run the function 1 million times not in parallel. We see in the below output that it took about 8 seconds to run not in parallel, while it only took about 4 seconds to run in parallel.
\newln
```{r}
# Not in parallel
start <- Sys.time() 
out <- replicate(1000000, rand.letter())
end <- Sys.time()
end - start # 8.139187

# In parallel
library(parallel)
start2 <- Sys.time() 
out2 <- unlist(mclapply(1:1000000, function(x) {rand.letter(x)}))
end2 <- Sys.time()
end2 - start2 # 4.274292
```

\newprob

## Problem 3 and 4

The link below is a link to my repo on GitHub called `tikitime`, which contains all of the files needed for the package `tikitime` that contains the function `rand.letter()`.  
\newln
https://github.com/cnkuhn/tikitime




